Relationships in the extremes and their influence on competition and coexistence

### 2021-10-01 paper/Paper.Rnw
s in abstract

\begin{abstract}
\noindent Understanding the coexistence of competing species is a topic of very broad 
interest in ecology. The modern coexistence
theory (MCT) of Chesson identifies several mechanisms of species coexistence,
one of which, storage effects, hinges on relationships between fluctuations
in environmental and competitive pressures on species populations. However, such relationships
have typically been measured using covariance methods, which do not account for 
the possibility that environmental and competitive influences may be more correlated
when these influences are strong than when they are weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATA) are extremely common
between ecological variables, and are important for a wide range of phenomena including 
extinction risk and the stability of ecosystem functioning. 
Here we extend the MCT to take into account the influence of ATAs. Storage
effects can be decomposed into two components, one of which quantifies the
influence of ATAs on species coexistence, and is a novel contribution of this study; and 
the other of which represents
the influence of environment-competition correlation \emph{per se}, without 
the influence of ATAs. Simulations of a simple model
%and an empirical example
illustrate that ATA influences on coexistence can be comparable in magnitude to other
mechanisms of coexistence, and that the presence or absence of ATAs can make the 
difference between species coexisting and competitive exclusion. 
\end{abstract}

>>>>

\begin{abstract}
\noindent Understanding the coexistence of competing species is a topic of very broad 
interest in ecology. The modern coexistence
theory (MCT) of Chesson identifies several mechanisms of species coexistence,
one of which, storage effects, hinges on relationships between fluctuations
in environmental and competitive pressures on species populations. However, such relationships
have typically been measured using covariance methods, which do not account for 
the possibility that environmental and competitive influences may be more correlated
when these influences are strong than when they are weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are extremely common
between ecological variables, and are important for a wide range of phenomena including 
extinction risk and the stability of ecosystem functioning. 
Here we extend the MCT to take into account the influence of ATAs. Storage
effects can be decomposed into two components, one of which quantifies the
influence of ATAs on species coexistence, and is a novel contribution of this study; and 
the other of which represents
the influence of environment-competition correlation \emph{per se}, without 
the influence of ATAs. Simulations of a simple model
%and an empirical example
illustrate that ATA influences on coexistence can be comparable in magnitude to other
mechanisms of coexistence, and that the presence or absence of ATAs can make the 
difference between species coexisting and competitive exclusion. 
\end{abstract}
